---
title: "jj_lab2"
author: "Jean-Luc Jackson"
date: \today
header-includes:
   - \usepackage{dcolumn}
output: 
  pdf_document: 
    number_sections: false
    toc: true
    toc_depth: 2
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning= FALSE)
knitr::knit_hooks$set(inline=function(x) {prettyNum(x, big.mark=",")})

if (!require("here")) install.packages("here")

library(here)
library(patchwork)
library(sandwich)
library(lmtest)
library(tidyverse)
library(stargazer)
```

```{r load data}
awesome_data <- read_rds(here ("datasets/awesome_data.rds"))
```

# Current Models
*Model Set 1 --- log models*

Model 1: log(truancy_rate) ~ log(pupil_teacher_fte_ratio)  \
Model 2: log(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility  \
Model 3: log(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility + average_salary  \

*Model Set 2 --- sqrt models*

Model 4: sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) \
Model 5: sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility \
Model 6: sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility + average_salary

*Model Set 3 --- sqrt interaction models*

Model 7: sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) \
Model 8: sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility \
Model 9: sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + average_salary

## Model Proposal
*Proposed Model Set 4 --- grade-level log interaction models*

Model 10: log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility \
Model 11: log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + grade_level \
Model 12: log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + grade_level + average_salary \


# Checking log(100*variable)

```{r}
df <- awesome_data$consolidated.school

df <- df %>% 
  mutate(
    truancy_rate = 100 * truancy_rate
  )

df %>% 
  ggplot() +
  aes(x = truancy_rate) +
  geom_histogram(bins = 20)

df %>% 
  ggplot() +
  aes(x = log(truancy_rate)) +
  geom_histogram(bins = 20)

model <- lm(log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + I(average_salary/1000),
            data = df)

summary(model)

```



# Adding Grade-Level Factor

```{r}
df_grade_levels <- awesome_data$schools

df_grade_levels <- df_grade_levels %>% 
  mutate(
    grade_level = case_when(
      x1st + x2nd + x3rd + x4th + x5th        == pk_12_count ~ "ELEM",
      x1st + x2nd + x3rd + x4th + x5th + x6th == pk_12_count ~ "ELEM",
      x7th + x8th                             == pk_12_count ~ "MIDD",
      x6th + x7th + x8th                      == pk_12_count ~ "MIDD",
      x9th + x10th + x11th + x12th            == pk_12_count ~ "HIGH",
      TRUE                                            ~ "MIX"
    )
  )

consolidated.school.grade_levels <- left_join(awesome_data$consolidated.school,
                df_grade_levels,
                by = c("county_code",
                       "county_name",
                       "district_code",
                       "district_name",
                       "school_code",
                       "school_name",
                       "high_school",
                       "high_school_plus"))

consolidated.school.grade_levels %>% 
  ggplot() +
  aes(x = grade_level) +
  geom_bar()

```

## Grade-Level Modeling

```{r}

df <- consolidated.school.grade_levels

# Re-ordering factors for optimal interpretability
df$grade_level <- factor(df$grade_level, levels=c('ELEM', 'MIDD', 'HIGH', 'MIX'))
df$lunch_subsidy_eligibility <- factor(df$lunch_subsidy_eligibility, levels = c('AVG', 'LOW', 'HIGH'))

model_10 <- lm(log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility,
               data = df)

model_11 <- lm(log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + grade_level,
               data = df)

model_12 <- lm(log(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + grade_level + average_salary,
               data = df)


# Add predicted column
df <- df %>% 
  mutate(
    model_pred_10 = predict(model_10),
    model_resid_10 = resid(model_10),
    model_pred_11 = predict(model_11),
    model_resid_11 = resid(model_11),
    model_pred_12 = predict(model_12),
    model_resid_12 = resid(model_12)
  )

# Plotting
# Model 10
df %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = log(truancy_rate),
      color = lunch_subsidy_eligibility) +
  geom_point(alpha = 0.5) +
  labs(title = "Model 10",
      x = "Pupil-Teacher Ratio (log scale)",
      y = "Truancy Rate (log scale)") +
  geom_line(size = 1,
            aes(y = model_pred_10))

# Model 11
df %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = log(truancy_rate),
      color = lunch_subsidy_eligibility,
      linetype = grade_level) +
  geom_point(alpha = 0.5) +
  labs(title = "Model 11",
      x = "Pupil-Teacher Ratio (log scale)",
      y = "Truancy Rate (log scale)") +
  geom_line(size = 1,
            aes(y = model_pred_11))

# Model 12
# can't be visualized because average_salary is a metric variable
df %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = log(truancy_rate),
      color = lunch_subsidy_eligibility,
      linetype = grade_level) +
  geom_point(alpha = 0.5) +
  labs(title = "Model 11",
      x = "Pupil-Teacher Ratio (log scale)",
      y = "Truancy Rate (log scale)") +
  geom_line(size = 1,
            aes(y = model_pred_11))

```

### Testing Coefficients

```{r}
mean(resid(model_10)^2)
mean(resid(model_11)^2)
mean(resid(model_12)^2)


coeftest(model_10, vcov = vcovHC)

coeftest(model_11, vcov = vcovHC)

coeftest(model_12, vcov = vcovHC)


waldtest(model_10, model_11, vcov=vcovHC(model_11, type="HC0"))

waldtest(model_11, model_12, vcov=vcovHC(model_12, type="HC0"))

```
```{r, results='asis', include=TRUE}

stargazer(model_10, model_11, model_12,
          title = "The Relationship Between Truancy and Pupil Student Ratio (Individual School)",
          column.labels=c("Model 10", "Model 11", "Model 12"),
          star.cutoffs = c(0.05, 0.01, 0.001),
          model.numbers = FALSE,
          omit.stat = c("f"),
          df = TRUE,
          single.row = FALSE,
          nobs = TRUE,
          align = TRUE,
          header = FALSE)

```


# Variable EDA
## Truancy Rate
### School-Level
```{r}
df <- awesome_data$consolidated.school

df %>% 
  ggplot() +
  aes(x = truancy_rate) +
  geom_histogram(bins = 50,
                 fill = "white",
                 color = "blue")

df %>% 
  ggplot() +
  aes(x = truancy_rate) +
  geom_histogram(bins = 50,
                 fill = "white",
                 color = "blue") +
  scale_x_continuous(trans = "log")

df %>% 
  ggplot() +
  aes(x = truancy_rate,
      fill = lunch_subsidy_eligibility) +
  geom_histogram(bins = 50)

df %>% 
  ggplot() +
  aes(x = truancy_rate,
      fill = lunch_subsidy_eligibility) +
  geom_histogram(bins = 50) +
  scale_x_continuous(trans = "log")

```


### District-Level
```{r}
df <- awesome_data$consolidated.district

df %>% 
  ggplot() +
  aes(x = truancy_rate) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue")

df %>% 
  ggplot() +
  aes(x = truancy_rate) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue") +
  scale_x_continuous(trans = "log")

df %>% 
  ggplot() +
  aes(x = sqrt(truancy_rate)) +
  geom_histogram(bins = 25)

```

## Pupil-Teacher Ratio
### School-Level

```{r}

df <- awesome_data$consolidated.school

df %>% 
  ggplot() +
  aes(x = pupil_teacher_fte_ratio) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue")

df %>% 
  ggplot() +
  aes(x = pupil_teacher_fte_ratio) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue") +
  scale_x_continuous(trans = "log")

df %>% 
  ggplot() +
  aes(x = pupil_teacher_fte_ratio,
      fill = lunch_subsidy_eligibility) +
  geom_histogram(bins = 25) +
  scale_x_continuous(trans = "log")
```

### District-Level

```{r}

df <- awesome_data$consolidated.district

df %>% 
  ggplot() +
  aes(x = pupil_teacher_fte_ratio) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue")

df %>% 
  ggplot() +
  aes(x = pupil_teacher_fte_ratio) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue") +
  scale_x_continuous(trans = "log")

df %>% 
  ggplot() +
  aes(x = pupil_teacher_fte_ratio,
      fill = lunch_subsidy_eligibility) +
  geom_histogram(bins = 25) +
  scale_x_continuous(trans = "log")
```

## Average Teacher Salary
### School-Level

```{r}

df <- awesome_data$consolidated.school

df %>% 
  ggplot() +
  aes(x = average_salary/1000) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue")

df %>% 
  ggplot() +
  aes(x = average_salary/1000,
      fill = charter) +
  geom_histogram(bins = 25)

df %>% 
  ggplot() +
  aes(x = average_salary/1000,
      fill = lunch_subsidy_eligibility) +
  geom_histogram(bins = 25)

```


### District-Level

```{r}

df <- awesome_data$consolidated.district

df %>% 
  ggplot() +
  aes(x = average_salary/1000) +
  geom_histogram(bins = 25,
                 fill = "white",
                 color = "blue")

df %>% 
  ggplot() +
  aes(x = average_salary/1000,
      fill = lunch_subsidy_eligibility) +
  geom_histogram(bins = 25)

```

## Grade Level
### School-Level
```{r}

df <- consolidated.school.grade_levels

hists <- df %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      fill = grade_level) +
  geom_histogram(bins = 40,
                 color = 'black') +
  geom_density()

dens <- df %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = ..density..,
      fill = grade_level) +
  geom_density(alpha = 0.5)

(hists / dens)
```

# Inspecting Models

```{r}
model_1 <- lm (log(truancy_rate) ~ log(pupil_teacher_fte_ratio), data=df)

model_2 <- lm (log(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility, data=df)

model_3 <- lm (log(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility + average_salary, data=df)

model_4 <- lm(sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio), data = df)

model_5 <- lm (sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility, data = df)

model_6 <- lm (sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility + average_salary, data = df)

model_7 <- lm(sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio), data = df)

model_8 <- lm (sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility, data = df)

model_9 <- lm (sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility + average_salary, data = df)

```


## School-Level

```{r}
# School-level data
df <- awesome_data$consolidated.school

# 3 Final Models
model_2 <- lm (log(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility,
               data = df)

model_5 <- lm (sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio) + lunch_subsidy_eligibility,
               data = df)

model_8 <- lm (sqrt(truancy_rate) ~ log(pupil_teacher_fte_ratio):lunch_subsidy_eligibility,
               data = df)



# Add predicted column
df_plot <- df %>% 
  mutate(
    model_pred_2 = predict(model_2),
    model_pred_5 = predict(model_5),
    model_pred_8 = predict(model_8)
  )

# Plotting
df_plot %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = log(truancy_rate),
      color = lunch_subsidy_eligibility) +
  geom_point(alpha = 0.5) +
  labs(title = "Model 2",
      x = "Pupil-Teacher Ratio (log scale)",
      y = "Truancy Rate (log scale)") +
  geom_line(size = 1,
            aes(y = model_pred_2))

df_plot %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = sqrt(truancy_rate),
      color = lunch_subsidy_eligibility) +
  geom_point(alpha = 0.5) +
  labs(title = "Model 5",
      x = "Pupil-Teacher Ratio (log scale)",
      y = "Truancy Rate (sqrt scale)") +
  geom_line(size = 1,
            aes(y = model_pred_5))

df_plot %>% 
  ggplot() +
  aes(x = log(pupil_teacher_fte_ratio),
      y = sqrt(truancy_rate),
      color = lunch_subsidy_eligibility) +
  geom_point(alpha = 0.5) +
  labs(title = "Model 8",
      x = "Pupil-Teacher Ratio (log scale)",
      y = "Truancy Rate (sqrt scale)") +
  geom_line(size = 1,
            aes(y = model_pred_8))

```




